The abundance of Bullet-Groups in $\Lambda$CDM

### 2014-03-18 paper/ms.tex
almost all comments are included

\title{The abundance of Bullet-Groups in $\Lambda$CDM}
\author{J. G. Fern\'andez-Trincado$^{1,2,3}$, J. E. Forero-Romero$^1$,
  T. Verdugo$^3$, G. Foex$^4$ and V. Motta$^4$} 
\affil{$^1$ Departamento de F\'{i}sica, Universidad de los Andes,
  Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\ 
       $^2$ Institute Utinam, CNRS UMR6213, Universit\'e de
  Franche-Comt\'e, OSU THETA de Franche-Comt\'e-Bourgogne,
  Besan\c{c}on, France\\ 
       $^3$ Centro de Investigaciones de Astronom\'ia, AP 264,
  M\'erida 5101-A, Venezuela\\
  $^4$ Departamento de F\'isica y Astronom\'ia, Universidad de
  Valpara\'iso, Avda. Gran Breta\~na 1111, Playa Ancha, Valpara\'iso
  2360102, Chile  
}

@@

We estimate the expected distribution of displacements between the two
dominant dark matter density peaks and between baryons and dark matter
in halos simulated within a full cosmological context. We use as a benchmark the
observation of \bullg, which is the lowest mass system observed so far
featuring a bi-modal dark matter distribution with a dislocated 
baryonic component. We extend previous results by
studying two halo samples: groups and clusters. We find that 50\% to 60\%
of the dark matter halos with circular velocities in the range
$300\kms$ to $700\kms$ (groups) show multimodal morphologies with
displacements between the dark matter clumps equal or larger than
$133\pm21$\hkpc as observed in \bullg. For dark matter halos with
circular velocities larger than $700\kms$ (clusters) this fraction
rises to $80\%$ to $85\%$. Using the same simulation we estimate the
dark matter-to-baryon spatial separation and find that $0.1\%$ to
$1.0\%$ of the groups should present separations equal or
larger than $87\pm 14$\hkpc corresponding to our observational
benchmark; for clusters this fraction is in the range of $4\%$ to
$10\%$, consistent with previous studies of dark matter to baryon
separations. The extension of the theoretical predictions and
observational results towards low mass bi-modal dark matter
configurations opens up the possibility for a new statistical test of
$\Lambda$CDM.   
\end{abstract}

@@

The Bullet Cluster (\bulla) provided a new
kind of observational evidence for the existence of dark matter
\citep{Markevitch2004,Clowe2006}. Quantifying the displacement between
dark matter and the dominant baryonic component (hot X-ray emitting gas)
has been used to test the Cold Dark Matter (CDM)
paradigm itself by quantifying the substructure velocity required to
produce such displacement \citep{Hayashi2006, Springel2007,
  Thompson2012}, the displacemeent between the dominant dark matter
and baryonic component \citep{ForeroRomero2010} to estimate the
expected abundance of such events in a $\Lambda$CDM Universe and even
explore possible extensions to the concordance cosmological model
\citep{Farrar2007,Lee2010,Lee2012}.

@@

However, a greater abundance of small mass systems has to be weighted
by the probability of having a merger and 
presenting a large displacement between the DM and baryonic
components. These two conditions (merger rates, maximum possible
displacement) are a funcion of halo mass in $\Lambda$CDM
cosmologies.  Such study has been performed for clusters but not for
lower mass systems \citep{ForeroRomero2010}. The existence of objects
like \bullg and observational campaigns like the Strong Lensing Legacy
Survey (SL2S) open up the possibility of finding multi-modal dark
matter distributions in the group mass range.

@@

This Letter is organized as follows. In Section
\ref{sec:simulation} we present the simulation and the halo catalogs
used in this work. We continue in Section \ref{sec:setup} with the
geometry of the problem at hand and the measurements setup. Next in
Section \ref{sec:results} we present our results and observational
perspectives to finally conclude in Sections \ref{sec:conclusions}.

@@

We use the Bolshoi Run, a cosmological DM only simulation over a cubic
volume of 250\hMpc comoving on a side. The simulation uses the ART code to
follow the evolution of a dark matter density field sampled with
$2048^3$ from $z=80$ to $z=0$. The cosmology used  corresponds
to  the spatially flat concordance model with the following
parameters:  the density parameter for matter (dark matter and baryons)
$\Omega_m=0.27$, the density parameter for baryonic matter
$\Omega_b=0.0469$, the density parameter for dark energy
$\Omega_{\Lambda}=0.73$, the Hubble parameter $h=0.7$, the
normalization of the Power spectrum $n=0.95$ and the amplitude of mass
density fluctuation (at redshift z$=$0) $\sigma_8=0.82$.  These
cosmological parameters are consistent with the nine-year Wilikinson
Microwave Anisotropy Probe (WMAP) results \citep{hinshaw_etal13}. A
detailed presentation of the simulation can be found in
\citet{2011ApJ...740..102K}.

The number of particles used for each of the DM component was
$2048^3$, resulting in a mass resolution of $1.35 \times 10^8$
M$_{\odot}$h$^{-1}$. The completeness limit in this simulation is set
for halos with $100$ particles corresponding to a mass of
$1.35\times10^{10}$\hMsun or a maximum circular velocity $V_{c}$ of
$50$\kms.

We use halo catalogs constructed using the Bound Density Maxima (BDM)
algorithm \citep{BDM,BDMb}. To define the radius of a halo we use a
density threshold of 360 times the mean density of the Universe, for 
different redshifts we use the overdensity criterion by
\cite{Bryan1998}. An important feature of BDM is that it allows us to
detect sub-halos inside larger virialized structures.

@@

In this Letter we make a study at four different redhifts $z=0.0,0.25,
0.5$ and $1.0$. First, we select all the host halos (i.e. halos that
are not inside a larger halo) with circular velocities $V_{\rm c}\geq
300$\kms. Then, we select the sub-halos with circular velocities
$V_{\rm c}\geq 75$\kms. The objective is to use this sub-halos as the
tracer of the sub-dominant dark matter clump in the merging cluster,
i.e. the bullets.

For the redshifts $z=0.0,0.25, 0.5$ and $1.0$ we find $10041, 10346,
10554$ and $10382$ host halos and $157853, 177331, 195072, 225188$
sub-halos, respectively. These two sets (host halos and sub-halos)
constitute the basis for our analysis. We find now for each host halo
we its most massive sub-halo. Each pair host halo/sub-halo is
considered as a potential Bullet system and is kept for the analysis
described in the next Section.

@@

The geometrical configuration can be further described by the
following quantities. The circular velocity of each component, $V_{\rm
  c,host}$ for the host and $V_{\rm c, sub}$ for the sub-halo and the
size of the host halo $R_{\rm vir}$. Another useful quantity computed
in the simulation is the distance between the minium of potential for
the host halo and its center of mass, $X_{\rm
  off}=||\vec{r}_{min}-\vec{r}_{cm}||/R_{\rm vir}$ which serves as a
measurement of how perturbed is the host halo.

As a zero-th order approximation, in this paper we work with three
quantities that are available from observations of Bullet-like
systems. The projected distance between two dominant DM clumps, the
projected distance between the DM-baryonic clumps and the ratio of the
mass associated to the DM clumps.

@@

In order to gain better insight, we use two
quantities that are not readily available from observations but can be
measured in the simulation. The first one is the sub-structure
velocity as a fraction of the host's circular velocity, $||\vec{v}||/V_{\rm c, host}$, as a measure of the the strength of the
merger.  The second is $\mu$ to measure the geometry (head-on
vs. grazing) of each collision.

The physical quantities described above can be used to describe
the three main stages in a bullet-like encounter. First, the sub-halo
crosses the virial radius of the host halo starting a head on
collision, $||\vec{r}||/R_{\rm vir}\approx 1$ and
$\mu\approx<0.0$. Second, as the sub-halo crosses for the first time
the center of the host halo $||\vec{r}||/R_{\rm vir} < 1.0$ and
$\mu>0.0$. Third, as the sub-halo reaches apogee and comes back to the
center of the halo $||\vec{r}||/R_{\rm vir} < 1.0$ and $\mu<0.0$. We use
this quantities in the enxt section to fully characterize the different
kind of interactions observed in the Bolshoi Run,

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{figure_1.eps}
\end{center}
\caption{
  Integrated probability distribution for the displacement
  between the center of the host halo and its dominant sub-halo. The left panel
  shows the results in terms of the physical displacements while
  the right panel shows the displacements normalized by the virial radius of
  the host halo. The continuous (dashed) line corresponds to the halos in the
  cluster (group).
  The vertical lines shows the mean and uncertainties in the estimated
  separation  between the two dark matter clumps in the results reported by
  \citet{Gastaldello} for the \bullg. Between $50\%$
  to $60\%$ of the groups show a displacement equal or larger than
  this observational benchmark.}
\label{fig:displacement}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{figure_2.eps} 
\end{center}
\caption{2D histogram in the plane $V_{\rm c,sub}/V_{\rm
    c,host}$-$d_{2D}$. The left panel corresponds to clusters and the
  right panel to groups. The star with error bars corresponds to
  \bullg data reported from \citet{2013A&A...552A..80M} and
  \citet{Gastaldello}. The data used to construct the histograms
  integrates the objects at all redshifts. }
\label{fig:mass_displacement}
\end{figure*}

@@

The panel with the projected 2D physical displacements also shows a
vertical stripe with the estimated displacement for the Bullet-group reported by
\cite{Gastaldello}. In the group sample we see that a fraction of
$50\%$ to $60\%$  should present a displacement equal than the
estimate for \bullg; in the cluster sample this fraction increases to
$80\%$-$85\%$. This fraction is naturally higher in more massive
systems because they are larger in size. Normalizing the displacements
by the virial radius, right panel Figure \ref{fig:displacement}, we
see that the distribution is the same regardless of the sample and the
redshift.

Figure \ref{fig:mass_displacement} shows 2D histograms in a plane
defined by the ratio of the two circular velocities $V_{\rm c,
  sub}/V_{\rm c, host}$ and the projected 2D physical displacements;
quantities that can be constrained by observations. The displacement
is a direct observable, while the ratio of the circular velocities can
be estimated from lensing studies or approximated by the ratio of the
total galaxy luminosity associated with the galaxy peaks or by a
lensing reconstruction.

To construct this figure we co-add all the halos in the sample (left,
clusters; right, groups) at all redshifts. We stack the data because
we do not observe any strong redshift dependence, additionally this allows
us to increase the signal in each bin. We overplot a star with error
bars that represents the observational estimates for the system
\bullg\ using the fraction in velocity dispersion in the line-of-sigth
of the group SL2S SJ08544-0121 ($\sigma_{host}=341^{+43}_{-109}$\kms and
$\sigma_{sub}=185^{+30}_{-62}$\kms) reported by
\citet{2013A&A...552A..80M} and the DM displacement inferred from the
data presented by \cite{Gastaldello}.

\subsection{Relative Velocities}
\label{sec:velocities}

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{figure_3.eps}
\end{center}
\caption{Integrated probability distribution for the relative velocity of the
sub-halo with respect to its host. The left panel shows the results
in physical units while the right panel show the same values
normalized by the circular velocity of the host halo. The line encoding
follows the same structure as Figure \ref{fig:displacement}}
\label{fig:velocities}
\end{figure*}

@@

In Figure \ref{fig:velocities} we present the integrated probability
of the relative peculiar velocities of the sub-halos with respect to
the host halo. The left panel presents this velocity in physical units
while the right panel presents them as a fraction of the circular
velocity of the host halo.

The panel with the normalized velocities shows that the distribution
of sub-halo velocities is close to universal. Regardless of the mass
of the host halo and the redshift the integrated distributions lie
very close to each other.  The universality of this profile extends to
the group mass range the findings of \citep{Hayashi2006} for clusters.

The median of this distribution is located at $v/V_{\rm
  c,host}=1.1$. We also note the strong break at $v/V_{\rm
  c,host}=3.0$ that is present in the data from the group sample that
allows us to probe fractions on the order of $10^{-4}$.  This break is
located close the escape velocity of $v/V_{\rm c,host}$ for dark
matter halos following a NFW profile with a concentration value
$c\approx 6$ \citep{Hayashi2006}.

\subsection{Collision Geometries}
\label{fig:geometry}

Figure \ref{fig:geometry} presents the geometry of the bullet groups using the 
variables $\mu$ and $D_{\rm off}$. The first evident feature is that
most of the configurations have $|mu|>0.9$ ($\theta\leq 30^{\circ}$),
meaning that most of the collisions can be described as a head-on
encounter while only a minority with $|mu|<0.9$ have grazing
trajectories. For the pairs on radial trajectories there are three
regions of interest in this plane that describe different stages in
the collision, assuming that the sub-halo merges (or falls below the
BDM detection threshold) right at its second pass through the center
of the host halo \citep{Poole2006}.

The first region has $\mu\approx-1$ and $D_{\rm off}>0.6$, which
locates the systems where a head-on collision has just started. The
sub-halo is close to the boundary of the host halo and is
infalling. The second region has $\mu\approx 1$ and $D_{\rm off}<0.6$;
at this stage the collision continues after the first crossing of the
host's center, the low number of halos with radial infalling velocities
and displacements $D_{\rm off}>0.6$ suggest that this the maximum
range of radii for the apogee.  The third region corresponds to
$\mu\approx-1$ and $D_{\rm off}<0.6$ which corresponds to the
secondary infall after apogee. We use this sequence in the next
subsection to estimate the expected displacement between baryons and
the dominant DM clump.

@@

Strictly speaking, the results we have derived so far, apply to multi-modal
groups and their expected separation between the two dominant dark
matter clumps. These displacements cannot be interpreted that all have a
corresponding DM-baryon displacement.

However, these different stages for a collision, that we describe at
the end of the previous sub-section, can produce different results in
terms of the displacement between dark matter and baryons. For
instance, the systems where the halo is starting to fall into the host
($\mu\approx-1$, $D_{\rm off}>0.6$) should not present a detectable
DM-baryon displacement.  We do expect such displacements when the
sub-structure has already passed through the center of the host halo,
i.e. cases where $\mu\approx 1$.

In this sub- section we estimate DM-baryon displacement statistics. We
work under the following hypothesis. First, we consider that systems
with $|\mu|<0.9$ have a baryonic displacement, $d_{2D}^{\rm bar}$,
equal to zero. This means that only head-on encounter produce a
displacement. Second, we consider that all systems with infalling
velocities $\mu<-0.9$ and large displacements $D_{\rm off}>0.6$ also
have baryonic displacements equal to zero. Third, for all the other
cases we estimate the displacement between the baryons and the
dominant DM peak by the distances between the dominant peak and the
center of mass of the main halo, $d_{2D}^{\rm bar} = X_{\rm off} R_{\rm
  vir}$, where $X_{\rm off}$ is the offset computed between the minium
of potential and the center of mass for each host halo.

@@

The results for the integrated distributions for $>d_{2D}^{\rm bar}$
are shown in Figure \ref{fig:baryonic_displacements}. The dashed lines
represented the results for groups and the continuous lines correspond
to clusters. As a test of our approach we compare the cluster results
against the analytic fit provided by \cite{ForeroRomero2010}. This fit
reproduces the statistics for the DM-baryon separation found for
clusters more massive than $>10^{14}$\hMsun in a simulation
that included a description for gas with $8$ times the volume of the
Bolshoi Simulation. The fit is valid for separations larger than
$70$\hkpc, beyond which we find that it provides a remarkably good
description within a factor of $\sim 2$ of our results. This gives us
confidence in our approach to estimate the expected fraction of
groups with a DM-baryon displacement.

@@

@@

In this Letter we estimate fraction of galaxy groups and clusters that
can present observational features associated to a bullet-like
event. This is motivated by the recent observational results of
\citep{Gastaldello} where a system (\bullg) on the mass range $1\times
10^{14}$\hMsun and velocity dispersion $650$\kms was reported to
feature a displacement between its baryonic (hot gas) and dark matter
components.

We estimate the distribution of projected displacements
between the dominant DM clumps in two kinds of systems; groups with
circular velocities $300\kms<V_{\rm c}<700\kms$ and clusters with $V_{\rm
  c}>700$\kms. We report these results at four different redshifts
$z=0.0,0.25,0.5$ and $10$. Our results based on large N-body dark
matter only cosmological simulation with such a resolution that allows
to study the group mass range for the first time in the context of
Bullet-like configurations.

Our main results is that that a fraction of $50\%$-$60\%$ of the halos
in the group sample present displacement equal or larger than the
observed displacement for \bullg. For halos in the cluster sample this
fraction increases to $80\%$-$85\%$. We also derive an
estimate for the displacement between the DM and the baryonic
component. In the group sample $0.1\%$-$1.0\%$ of the halos show a
displacement equal or larger than  the  measurements of \bullg by
\citep{Gastaldello}. In the cluster sample this fraction rises to
$4\%$-$10\%$.

>>>>

\title{The abundance of Bullet-Groups in $\Lambda$CDM}
\author{J. G. Fern\'andez-Trincado$^{1,2,3}$, J. E. Forero-Romero$^1$,
  T. Verdugo$^3$, G. Foex$^4$ and V. Motta$^4$} 
\affil{$^1$ Departamento de F\'{i}sica, Universidad de los Andes,
  Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\ 
       $^2$ Institute Utinam, CNRS UMR6213, Universit\'e de
  Franche-Comt\'e, OSU THETA de Franche-Comt\'e-Bourgogne,
  Besan\c{c}on, France\\ 
       $^3$ Centro de Investigaciones de Astronom\'ia, AP 264,
  M\'erida 5101-A, Venezuela\\
  $^4$ Instituto de F\'isica y Astronom\'ia, Universidad de
  Valpara\'iso, Avda. Gran Breta\~na 1111, Playa Ancha, Valpara\'iso
  2360102, Chile  
}

@@

We estimate the expected distribution of displacements between the two
dominant dark matter density peaks and between baryons and dark matter
in halos simulated within a full cosmological context. We use as a benchmark the
observation of \bullg, which is the lowest mass system observed so far
featuring a bi-modal dark matter distribution with a dislocated 
baryonic component. We extend previous results by
studying two halo samples: groups and clusters. We find that $(55\pm 5)\%$
of the dark matter halos with circular velocities in the range
$300\kms$ to $700\kms$ (groups) show multimodal morphologies with
displacements between the dark matter clumps equal or larger than
$133\pm21$\hkpc as observed in \bullg. For dark matter halos with
circular velocities larger than $700\kms$ (clusters) this fraction
rises to $(82\pm 3)\%$. Using the same simulation we estimate the
dark matter-to-baryon spatial separation and find that $0.1$ to
$1.0\%$ of the groups should present separations equal or
larger than $87\pm 14$\hkpc corresponding to our observational
benchmark; for clusters this fraction rises to $(7\pm 3)\%$, consistent
with previous studies of dark matter to baryon separations. The
extension of the theoretical predictions and observational results
towards low mass bi-modal dark matter configurations opens up the
possibility for a new statistical test of  $\Lambda$CDM.
\end{abstract}

@@

The Bullet Cluster (\bulla) provided a new kind of observational
evidence for the existence of dark matter
\citep{Markevitch2004,Clowe2006}. Quantifying the displacement between
dark matter and the dominant baryonic component (hot X-ray emitting gas)
has been used to test the Cold Dark Matter (CDM)
paradigm itself by quantifying the substructure velocity required to
produce such displacement \citep{Hayashi2006, Springel2007,
  Thompson2012}, the displacemeent between the dominant dark matter
and baryonic component \citep{ForeroRomero2010} to estimate the
expected abundance of such events in a $\Lambda$CDM Universe, putting
constraints on the dark matter particle self-interaction cross section
and even explore possible extensions to the concordance cosmological model
\citep{Farrar2007,Lee2010,Lee2012}.

@@

However, a greater abundance of small mass systems has to be weighted
by the probability of having a merger and 
presenting a large displacement between the DM and baryonic
components. These two conditions (merger rates, maximum possible
displacement) are a funcion of DM halo mass in $\Lambda$CDM
cosmologies.  Such study has been performed for clusters but not for
lower mass systems \citep{ForeroRomero2010}. The existence of objects
like \bullg\ and observational campaigns like the Strong Lensing Legacy
Survey (SL2S) \citep{Cabanac2007,More2012} open up the possibility of finding
multi-modal dark matter distributions in the group mass range.

@@

This Letter is organized as follows. In Section
\ref{sec:simulation} we present the simulation and the halo catalogs
used in this work. We continue in Section \ref{sec:setup} with the
geometry of the problem at hand and the measurements setup. Next in
Section \ref{sec:results} we present our results and observational
perspectives to finally conclude in Section \ref{sec:conclusions}.

@@

We use the Bolshoi Run, a cosmological DM only simulation over a cubic
volume of 250\hMpc comoving on a side. The simulation uses the ART
code \citep{Kravtsov1997} to follow the evolution of a dark matter
density field sampled with $2048^3$ from $z=80$ to $z=0$. The
cosmology used  corresponds to  the spatially flat concordance model
with the following parameters:  the density parameter for matter (dark
matter and baryons) $\Omega_m=0.27$, the density parameter for baryonic matter
$\Omega_b=0.0469$, the density parameter for dark energy
$\Omega_{\Lambda}=0.73$, the Hubble parameter $h=0.7$, the
normalization of the Power spectrum $n=0.95$ and the amplitude of mass
density fluctuation (at redshift z$=$0) $\sigma_8=0.82$.  These
cosmological parameters are consistent with the nine-year Wilikinson
Microwave Anisotropy Probe (WMAP) results \citep{hinshaw_etal13}. A
detailed presentation of the simulation can be found in
\citet{2011ApJ...740..102K}.

The number of particles used for each of the DM component was
$2048^3$, resulting in a mass resolution of $1.35 \times 10^8$
M$_{\odot}$h$^{-1}$. The completeness limit in this simulation is set
for halos with $100$ particles corresponding to a mass of
$1.35\times10^{10}$\hMsun\ or a maximum circular velocity $V_{c}$ of
$50$\kms.

We use halo catalogs constructed using the Bound Density Maxima (BDM)
algorithm \citep{BDM,BDMb}. To define the radius of a halo we use a
density threshold of 360 times the mean density of the Universe (or
200 times the critical density), for different redshifts we use the
overdensity criterion by \cite{Bryan1998}. An important feature of BDM
is that it allows us to detect sub-halos inside larger virialized
structures.

@@

In this Letter we make a study at four different redshifts $z=0.0,0.25,
0.5$ and $1.0$. First, we select all the host halos (i.e. halos that
are not inside a larger halo) with circular velocities $V_{\rm c}\geq
300$\kms ($\geq 1\times 10^{13}\\hMsun$). Then, we select the sub-halos with
circular velocities $V_{\rm c}\geq 75$\kms ($\geq 5\times
10^{10}\hMsun$). The objective is to use this sub-halos as the tracer
of the sub-dominant dark matter clump in the merging cluster, i.e. the
bullets.

For the redshifts $z=0.0,0.25, 0.5$ and $1.0$ we find $10041, 10346,
10554$ and $10382$ host halos and $157853, 177331, 195072, 225188$
sub-halos, respectively. These two sets (host halos and sub-halos)
constitute the basis for our analysis. We associate each host halo
to its most massive sub-halo, each one of these pairs (host \& sub-halo)
is considered as a potential Bullet system and is kept for the
analysis described in the next Section.

@@

The geometrical configuration can be further described by the
following quantities. The circular velocity of each component, $V_{\rm
  c,host}$ for the host and $V_{\rm c, sub}$ for the sub-halo and the
size of the host halo $R_{\rm vir}$. Another useful quantity computed
in the simulation is the distance between the minium of potential for
the host halo and its center of mass, $X_{\rm
  off}=||\vec{r}_{min}-\vec{r}_{cm}||/R_{\rm vir}$ which serves as a
measurement of how much the host halo is perturbed.

As a zero-th order approximation, in this paper we work with three
observed quantities for Bullet-like systems. The projected distance
between two dominant DM clumps, the projected distance between the
DM-baryonic clumps and the ratio of the mass associated to the DM
clumps.

@@

%In order to gain better insight, we use two
%quantities that can be infered from the simulation. The first one is
%the sub-structure velocity as a fraction of the host's circular
%velocity, $||\vec{v}||/V_{\rm c, host}$, as a measure of the the
%strength of the merger.  The second onw is $\mu$, which measures the geometry
%(head-on vs. grazing) of each interaction.

The physical quantities described above can be used to describe
the three main stages in a bullet-like encounter. First, the sub-halo
crosses the virial radius of the host halo starting a head on
collision, $||\vec{r}||/R_{\rm vir}\approx 1$ and
$\mu\lessapprox<0.0$. Second, as the sub-halo crosses for the first time
the center of the host halo $||\vec{r}||/R_{\rm vir} < 1.0$ and
$\mu>0.0$. Third, as the sub-halo reaches apogee and comes back to the
center of the halo $||\vec{r}||/R_{\rm vir} < 1.0$ and $\mu<0.0$. We use
these quantities in the next section to fully characterize the different
kind of interactions observed in the Bolshoi Run,

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{figure_1.eps}
\end{center}
\caption{
  Integrated probability distribution for the displacement between the
  center of the host halo and its dominant sub-halo. The left panel
  shows the results in terms of the physical displacements while the
  right panel shows the displacements normalized by the virial radius
  of the host halo. The continuous (dashed) line corresponds to the halos in the
  cluster (group).
  The vertical lines show the mean value and uncertainties ($133\pm
  21$\hkpc) in the  separation between the two dark matter clumps
  estimated in \citet{Gastaldello} for the \bullg. Between $50\%$
  to $60\%$ of the groups show a displacement equal or larger than
  this observational benchmark.}
\label{fig:displacement}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{figure_2.eps} 
\end{center}
\caption{2D histogram in the plane $V_{\rm c,sub}/V_{\rm
    c,host}$-$d_{2D}$. The left panel corresponds to clusters and the
  right panel to groups. The star with error bars corresponds to
  \bullg\ data reported from \citet{2013A&A...552A..80M} and
  \citet{Gastaldello}. The data used to construct the histograms
  integrates the objects at all redshifts. }
\label{fig:mass_displacement}
\end{figure*}

@@

The panel with the projected 2D physical displacements also shows a
vertical stripe with the estimated displacement for the Bullet-group reported by
\cite{Gastaldello}. In the group sample we see that a fraction of
$50\%$ to $60\%$  should present a displacement equal than the
estimate for \bullg; in the cluster sample this fraction increases to
$80\%$-$85\%$. The uncertainties in this estimates are derived from
the uncertainties in the displacement measurement for \bullg.This
fraction is naturally higher in more massive systems because they are
larger in size. Normalizing the displacements by the virial radius,
right panel Figure \ref{fig:displacement}, we see that the
distribution is the same regardless of the sample and the redshift.

Figure \ref{fig:mass_displacement} shows 2D histograms in a plane
defined by the ratio of the two circular velocities $V_{\rm c,
  sub}/V_{\rm c, host}$ and the projected 2D physical displacements;
quantities that can be constrained by observations.

To construct this figure we co-add all the halos in the sample (left,
clusters; right, groups) at all redshifts. We stack the data because
we do not observe any strong redshift dependence. This allows
us to increase the signal in each bin. We overplot a star with error
bars that represents the observational constraints for the system
\bullg\ using the fraction in velocity dispersion in the line-of-sigth
of the group SL2S SJ08544-0121 ($\sigma_{host}=341^{+43}_{-109}$\kms and
$\sigma_{sub}=185^{+30}_{-62}$\kms) reported by
\citet{2013A&A...552A..80M} and the DM displacement inferred from the
data presented by \cite{Gastaldello}.

%\subsection{Relative Velocities}
%\label{sec:velocities}

%\begin{figure*}
%\begin{center}
%\includegraphics[width=0.9\textwidth]{figure_3.eps}
%\end{center}
%\caption{Integrated probability distribution for the relative velocity of the
%sub-halo with respect to its host. The left panel shows the results
%in physical units while the right panel show the same values
%normalized by the circular velocity of the host halo. The line encoding
%follows the same structure as Figure \ref{fig:displacement}}
%\label{fig:velocities}
%\end{figure*}

@@

%In Figure \ref{fig:velocities} we present the integrated probability
%of the relative peculiar velocities of the sub-halos with respect to
%the host halo. This velocity is presented in the left panel in
%physical units and in the right panel as a fraction of the circular
%velocity of the host halo.

%The panel with the normalized velocities shows that the distribution
%of sub-halo velocities is close to universal. Regardless of the mass
%of the host halo and the redshift the integrated distributions lie
%very close to each other.  The universality of this profile also holds
%for the group mass range. extending the result found by
%\citep{Hayashi2006} for cluster mass halos.

%The median of this distribution is located at $v/V_{\rm
%  c,host}=1.1$. We also note the strong break at $v/V_{\rm
%  c,host}=3.0$ present in the data from the group sample, which
%allows us to probe fractions on the order of $10^{-4}$.  This break is
%located close the escape velocity of $v/V_{\rm c,host}$ for dark
%matter halos following a NFW profile with a concentration value
%$c\approx 6$ \citep{Hayashi2006}.

\subsection{Collision Geometries}
\label{sec:geometry}

Figure \ref{fig:geometry} presents the geometry of the bullet groups using the 
variables $\mu$ and $D_{\rm off}$. The first evident feature is that
most of the configurations have $|\mu|>0.9$ ($\theta\leq 30^{\circ}$),
meaning that most of the collisions can be described as a head-on
encounter while only a minority with $|\mu|<0.9$ have grazing
trajectories. For the pairs on radial trajectories there are three
regions of interest in this plane that describe different stages in
the collision, assuming that the sub-halo merges (or falls below the
BDM detection threshold) right at its second pass through the center
of the host halo \citep{Poole2006}.

The first region has $\mu\approx-1$ and $D_{\rm off}>0.6$, which
locates the systems where a head-on collision has just started. The
sub-halo is close to the boundary of the host halo and is
infalling. The second region has $\mu\approx 1$ and $D_{\rm off}<0.6$;
at this stage the collision continues after the first crossing of the
host's center, the low number of halos with radial infalling velocities
and displacements $D_{\rm off}>0.6$ suggest that this the maximum
range of radii for the apogee.  The third region corresponds to
$\mu\approx-1$ and $D_{\rm off}<0.6$ that is the the
secondary infall after apogee. We use this sequence in the next
subsection to estimate the expected displacement between baryons and
the dominant DM clump.

@@

Strictly speaking, the results we have derived so far apply to multi-modal
groups and their expected separation between the two dominant dark
matter clumps. However, a non-zero DM-DM displacement cannot be
interpreted as having a corresponding non-zero DM-baryon
displacement. Given that we use a DM only simulation, we have to find
a way to estimate DM-baryon displacements from the available information.

This estimate is based on the different collision stages described at
the end of the previous sub-section.  For instance, in systems where
the halo is starting to fall into the host ($\mu\approx-1$, $D_{\rm
  off}>0.6$) the DM-baryon displacement should be equal to
zero. However, when the  sub-structure has already passed through the
center of the host halo, i.e. cases where $\mu\approx 1$, the
DM-baryon displacement should be different from zero.

To estimate the DM-baryon displacement we  work under the following
hypothesis. First, we consider that systems with $|\mu|<0.9$ have a
baryonic displacement, $d_{2D}^{\rm bar}$, equal to zero. This means
that only head-on encounters produce a displacement. Second, we
consider that all systems with infalling velocities $\mu<-0.9$ and
large displacements $D_{\rm off}>0.6$ also have baryonic displacements
equal to zero. Third, for all the other cases we estimate the
displacement between the baryons and the dominant DM peak by
$d_{2D}^{\rm bar} = X_{\rm off} R_{\rm   vir}$, where $X_{\rm   off}$
is the offset computed between the minium of potential and the center
of mass for each host halo computed from all the matter inside the
virial radius.

@@

The results for the integrated distributions for $>d_{2D}^{\rm bar}$
are shown in Figure \ref{fig:baryonic_displacements}. The dashed lines
represented the results for groups and the continuous lines correspond
to clusters. As a test of our model we compare the cluster results
against the analytic fit provided by \cite{ForeroRomero2010}. This fit
reproduces the statistics for the DM-baryon separation found for
clusters more massive than $>10^{14}$\hMsun\ in a simulation
which included a description for DM and gas with $8$ times the volume of the
Bolshoi Simulation. The fit is valid for separations larger than
$70$\hkpc, beyond which we find that it provides a remarkably good
description within a factor of $\sim 2$ of our results. This gives us
confidence in our approach to estimate the expected fraction of
groups with a DM-baryon displacement.

@@

\subsection{Possible Observational Statistical Tests}

Recently \citep{Foex2013} presented an analysis of $80$ galaxy groups
from the SL2S sample. From the light distribution, only $34$ objects
($\approx 42\%$) have regular isophotes, $33$ had ellongated
isophotes (hints of merging system), and $13$ ($\approx 16\%$) had a
clear bi-modal light distribution.

The bi-modal objects are defined to have at least a clear second
luminosity peak within $350$\hkpc from the main halo, as traced by the
strong lensing system.  The lowest separation in those systems is
$64$ \hkpc and the average is $145\pm 52$\hkpc. Comparing this average
displacement against the integratied distribution for the group sample (left
panel, Figure \ref{fig:displacement}) one would expects between $40\%$
to $80\%$ of DM halos to present those displacements. This fraction is
certainly larger than the $\approx 16\%$ inferred from the
observational sample.

We have taken the results from the simulation and selected all the
objects in the group sample with a DM-DM displacement less than
$350$\hkpc but larger than $64$\hkpc to find XXX objects in
total. From this subsample we find that XXX have a DM-baryon
displacement larger than $87$\hkpc.

This rough comparison hints that it is possible to find a large fraction
of multi-modal groups with a DM-baryon displacement in a $\Lambda$CDM
cosmology. A proper comparison must take into account all the
observational uncertainties and biases to derive a stronger prediction from the
simulation, something that is beyond the scope of this Letter.

@@

In this Letter we estimate fraction of galaxy groups and clusters that
can present observational features associated to a bullet-like
event. This is motivated by the recent observational results of
\citep{Gastaldello} where a system (\bullg) on the mass range $1\times
10^{14}$\hMsun\ and velocity dispersion $650$\kms was reported to
feature a displacement between its baryonic (hot gas) and dark matter
components.

We estimate the distribution of projected displacements
between the dominant DM clumps in two kinds of systems; groups with
circular velocities $300\kms<V_{\rm c}<700\kms$ ($1.0\times
10^{13}\hMsun < M_{\rm vir}<1.0\times 10^{14}\hMsun$) and clusters with $V_{\rm
  c}>700$\kms ($M_{\rm vir}>1.0\times 10^{14}\hMsun$). We report these
results at four different redshifts $z=0.0,0.25,0.5$ and $10$. Our
results are based on large DM-only N-body cosmological simulation with
a resolution that allows to study for the first time Bullet-like
configurations in the mass range of galaxy groups.

Our main results is that that a fraction of $50\%$-$60\%$ of the halos
in the group sample present displacement equal or larger than the
observed displacement for \bullg. For halos in the cluster sample this
fraction increases to $80\%$-$85\%$. We also derive an
estimate for the displacement between the DM and the baryonic
component. In the group sample $0.1\%$-$1.0\%$ of the halos show a
displacement equal or larger than  the  measurements of \bullg\ by
\citep{Gastaldello}. In the cluster sample this fraction rises to
$4\%$-$10\%$.
